\sffamily An EEG-based visual Brain-Computer Interface for gaze-free communication

### 2024-10-06 chapters/bci.tex
cover page

\todo{comments liesel paradigms}
\todo{spellcheck}
\todo{are all acronyms defined}
\todo{fix all latex warnings to make sure no refs are missing (or grep on that
specific warning)}
\todo{uniform ref or autoref}
\todo{don't capitalize figure, table, algorithm,...}
\todo{define short captions for tables and figures}
\todo{fix lacheck/chktex notifications}
\todo{use terminology individuals with severe
speech and physical impairments (SSPI), avoid phrasing 'suffering from'}
\todohsm{more subtitles}
\todofc{double check statements about what has been published 'for the sake of clarity, the introduction has been mofed to"}
\todofc{chapters should look more like full papers}
\todo{uniform vsa condition vsa setting}
\todo{acronyms written in full every new chapter and in captions}
\todo{search for ??}
\todo{acro usage in chapters}

People with \ac{sspi} due to
neurodegenenerative disease, stroke, traumatic brain or spine injury
become entirely reliant on caregivers or family members in their day to day
routine.
Wouldn't it be great if we could help them regain some quality of life by
making them more independent?

@@

\subsubsection{\Acf{lda}}
\Ac{lda} is a more well-known method that is, int fact, nearly equivalent to
\ac{lcmv} beamforming~\cite{Treder2016} under certain conditions.
It assumes normally distributed data with a different mean per class but a
common covariance.
Similar to \ac{lcmv}-beamforming, \ac{lda} learns a weight vector $\mat{w}$
that satisfies
\begin{equation}
  \argmax_\mat{w}\frac{\mat{w}^\intercal\mat{S}_\text{b}\mat{w}}{\mat{w}^\intercal\mat{S}_\text{w}\mat{w}}
\end{equation}
The between-class scatter matrix $\mat{S}_\text{b}$ models the variability
between the expected responses, the within-class scatter matrix
$\mat{S}_\text{w}$ the expected noise.
The optimization criterion corresponds to finding a projection that maximizes
the spread between classes (signal) while minimizing the spread within classes
(noise).
$mat{w}$ can be obtained as eigenvector of
$\mat{S}_\text{w}^{-1}\mat{S}_\text{b}$ corresponding to the largest eigenvalue.

@@

Another example of this is \ac{tlda}~\cite{Sosulski2022}, which assumes temporal
stationarity on the background noise to regularize the problem.
It does this by imposing a block-Toeplitz structure to the between-class covariance
matrix of \ac{lda}, such that every temporal covariance block corresponding to
a channel has the structure
\begin{equation}
  \mat{S}_\text{w}^(c) = \begin{bmatrix}
s_0 & s_1 & s_2 & \cdots & s_{k-1} \\
s_1 & s_0 & s_1 & \cdots & s_{k-2} \\
s_2 & s_1 & s_0 & \cdots & s_{k-3} \\
\vdots & \vdots & \vdots & \ddots & \vdots \\
s_{k-1} & s_{k-2} & s_{k-3} & \cdots & s_0
\end{bmatrix}
\end{equation}
This method is simple yet effective reaching state-of-the-art decoding
performance.

>>>>

\todo{comments liesel paradigms}
\todo{spellcheck}
\todo{are all acronyms defined}
\todo{fix all latex warnings to make sure no refs are missing (or grep on that
specific warning)}
\todo{uniform ref or autoref}
\todo{don't capitalize figure, table, algorithm,...}
\todo{define short captions for tables and figures}
\todo{fix lacheck/chktex notifications}
\todo{use terminology individuals with severe
speech and physical impairments (SSPI), avoid phrasing 'suffering from'}
\todohsm{more subtitles}
\todofc{double check statements about what has been published 'for the sake of clarity, the introduction has been mofed to"}
\todofc{chapters should look more like full papers}
\todo{uniform vsa condition vsa setting}
\todo{acronyms written in full every new chapter and in captions}
\todo{search for ??}
\todo{acro usage in chapters}
\chapter{Brain-computer interfaces}
\label{sec:bci}
People with \ac{sspi} due to
neurodegenenerative disease, stroke, traumatic brain or spine injury
become entirely reliant on caregivers or family members in their day to day
routine.
Wouldn't it be great if we could help them regain some quality of life by
making them more independent?

@@

\subsubsection{\Acf{lda}}
\Ac{lda} is a more well-known method that is, in fact, nearly equivalent to
\ac{lcmv} beamforming~\cite{Treder2016} under certain conditions.
It assumes normally distributed data with a different mean per class but a
common covariance.
Similar to \ac{lcmv}-beamforming, \ac{lda} learns a weight vector $\mat{w}$
that satisfies
\begin{equation}
  \argmax_\mat{w}\frac{\mat{w}^\intercal\mat{S}_\text{b}\mat{w}}{\mat{w}^\intercal\mat{S}_\text{w}\mat{w}}
\end{equation}
The between-class scatter matrix $\mat{S}_\text{b}$ models the variability
between the expected responses, the within-class scatter matrix
$\mat{S}_\text{w}$ the expected noise.
The optimization criterion corresponds to finding a projection that maximizes
the spread between classes (signal) while minimizing the spread within classes
(noise).
$\mat{w}$ can be obtained as eigenvector of
$\mat{S}_\text{w}^{-1}\mat{S}_\text{b}$ corresponding to the largest eigenvalue.

@@

Another example of this is \ac{tlda}~\cite{Sosulski2022}, which assumes temporal
stationarity on the background noise to regularize the problem.
It does this by imposing a block-Toeplitz structure to the between-class covariance
matrix of \ac{lda}, such that every temporal covariance block corresponding to
a channel has the structure
\begin{equation}
  \mat{S}_\text{w}^{(c)} = \begin{bmatrix}
s_0 & s_1 & s_2 & \cdots & s_{k-1} \\
s_1 & s_0 & s_1 & \cdots & s_{k-2} \\
s_2 & s_1 & s_0 & \cdots & s_{k-3} \\
\vdots & \vdots & \vdots & \ddots & \vdots \\
s_{k-1} & s_{k-2} & s_{k-3} & \cdots & s_0
\end{bmatrix}
\end{equation}
This method is simple yet effective reaching state-of-the-art decoding
performance.
